## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(BAS)
library(MASS)
library(GGally)
library(gridExtra)
```

### Load data

```{r load-data}
load("movies.Rdata")
```

### Part 1: Data

The dataset given for this project consists of information about the movies released between the years 1970–2016. The source of the information is taken from Rotten Tomatoes and IMDB websites.

The dataset have 651 random samples (movies) with 32 available variables in it. The dataset provided an overall picture about the type of movie, run time, date and year of release of the movie, IMDB rating and so many other factors.

The dataset also provide information about the views and opinion of the audiences and critics. The given dataset is an observational study only and based on the data, the success or failure of the movie could be identified. But there may be so many other reasons like story, music, central message and other aspects which may contribute to the success of the movie. From the dataset, we can’t estimate other factors that may lead to popularity of the movies. So, therefore there are no casual inferences from the data. But we can observe the features and interest of the audiences toward the movies that has changed from 1970 to 2016.


### Part 2: Data manipulation


Here We are first going to create some new variables using the mutate function to support in our exploratory data analysis.
The summary of the new variables are below.

feature_film: "yes" if ```title_type``` is Feature Film, "no" otherwise.
drama: "yes" if ```genre``` is Drama, "no" otherwise.
mpaa_rating_R: "yes" if ```mpaa_rating``` is R, "no" otherwise
oscar_season: "yes" if movie is released in November, October, or December (based on ```thtr_rel_month```), "no" otherwise.
summer_season: "yes" if movie is released in May, June, July, or August (based on ```thtr_rel_month```), "no" otherwise.




```{r}
movies <- movies %>%
  mutate(feature_film = ifelse(title_type == "Feature Film", "yes", "no"),
         drama = ifelse(genre == "Drama", "yes", "no"),
         mpaa_rating_R = ifelse(mpaa_rating == "R","yes","no"),
         oscar_season = ifelse(thtr_rel_month == 11 | thtr_rel_month == 10 | thtr_rel_month == 12, "yes", "no"),
         summer_season = ifelse(thtr_rel_month == 5 | thtr_rel_month == 6 | thtr_rel_month == 7 | thtr_rel_month == 8, "yes","no"))
```

Now we are going to create a new dataframe called " movies2 ". These dataframe includes all the subset of the total variables.


```{r}
movies2_features <- c("audience_score", "feature_film", "drama", "runtime", "mpaa_rating_R", "thtr_rel_year", "oscar_season", "summer_season", "imdb_rating", "imdb_num_votes", "critics_score", "best_pic_nom", "best_pic_win", "best_actor_win", "best_actress_win", "best_dir_win", "top200_box")
movies2 <- movies[movies2_features]
```


### Part 3: Exploratory data analysis


```{r}
summary(movies2)
```

First let us see the summary of the dataframe, it gives us the summary of lay out of each variable.

Now, Lets have a look at the levels of each variable.

```{r}
str(movies2)
```

Let u see how the newly formed variables interact with audience score using the boxplot.

```{r}
plot1 <- ggplot(movies2, aes(x=mpaa_rating_R,y=audience_score))+
            geom_boxplot(colour="aquamarine4")
  
plot2 <- ggplot(movies2, aes(x=oscar_season, y=audience_score))+
            geom_boxplot(colour="aquamarine4")
  
plot3 <- ggplot(movies2, aes(x=summer_season,y=audience_score))+
            geom_boxplot(colour="aquamarine4")
  
plot4 <- ggplot(movies2, aes(x=feature_film, y=audience_score))+
            geom_boxplot(colour="aquamarine4")
  
plot5 <- ggplot(movies2, aes(x=drama, y=audience_score))+
            geom_boxplot(colour="aquamarine4")
            
grid.arrange(plot1,plot2,plot3,plot4,plot5, ncol=3)

```


Now lets map out the correlation charts between audience score and all the other variables in the dataframe  movies2. This correlation charts will show the relation between them.

```{r, fig.width = 10, fig.height = 10}
suppressWarnings(suppressMessages(print(ggpairs(movies2, columns = 1:8))))
```

```{r, fig.width = 10, fig.height = 10}
suppressWarnings(suppressMessages(print(ggpairs(movies2, columns = c(1,9:17)))))
```


From the correlation chart we can observe that there is high correlation between audience_score with critics_score and imdb_rating of the dataframe movies2.


```{r}
cor(movies2$audience_score, movies2$critics_score)
```

The correlation between audience_score and critics_score is 0.7042762.


```{r}
cor(movies2$audience_score, movies2$imdb_rating)
```


The correlation between audience_score and imdb_rating is 0.8648652.


Now let us visualize this correlation of both critics_score and imdb_rating with audience_score using a scatter plot with a regression line.

```{r}
ggplot(data=movies2, aes(x = audience_score, y = critics_score)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```


```{r}
ggplot(data=movies2, aes(x = audience_score, y = imdb_rating)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```


We can observe there is positive correlations for the both sets of variables.


### Part 4: Modeling

Firstly create the full linear model, incorporating every variables in the dataframe movies2.

Will use the stepAIC function from library MASS to build a model (backwards) until the AIC can not be lowered.

```{r}
as_full <- lm(audience_score ~ ., data= na.omit(movies2))
as_full
```

### Creating a model based on AIC

We will use the stepAIC function, tuned to optimize for AIC, to find the best model. The model will be built backwards.

```{r}
stepAIC.model <- stepAIC(as_full, direction = "backward", trace = TRUE)
```

The final model built using AIC consists of the following variables:

```runtime + mpaa_rating_R + thtr_rel_year + imdb_rating + critics_score + best_pic_nom + best_actor_win```
```{r}
AIC.lm <- lm(audience_score ~ runtime + mpaa_rating_R + thtr_rel_year + imdb_rating + critics_score + best_pic_nom + best_actor_win + best_actress_win, data=movies2)
```

Having a look at the coefficients of this model:

```{r}
AIC.lm$coefficients
```

Having a look at the standard deviation of the model:

```{r}
summary(AIC.lm)$sigma
```

```{r}
#plot(movies2$audience_score ~ AIC.lm$residuals)
```

Plotting the residuals of the model:

```{r}
ggplot(data=AIC.lm, aes(x=AIC.lm$residuals)) + geom_histogram()
```

We can observe that there the residuals are normally distributed. 

### Creating a model based on BIC

We will use the stepAIC function, tuned to optimize for BIC, to find the best model. The model will be built backwards.


```{r}
stepBIC.model <- stepAIC(as_full, direction = "backward", k=log(nrow(movies2)), trace = TRUE)
```

The final model will use the following variables:

```audience_score ~ runtime + imdb_rating + critics_score```
```{r}
BIC.lm <- lm(audience_score ~ runtime + imdb_rating + critics_score, data=movies2)
```

```{r}
BIC.lm$coefficients
```

```{r}
summary(BIC.lm)$sigma
```

```{r}
#plot(na.omit(movies2$audience_score) ~ BIC.lm$residuals)
```

Taking a look at the residuals:

```{r}
ggplot(data=BIC.lm, aes(x=BIC.lm$residuals)) + geom_histogram()
```

We can see that the residuals are normally distributed. 


### Creating a model using bayesian averaging

```{r}
as_full.bas <- bas.lm(audience_score ~ .,
       prior ="BIC",
       modelprior = uniform(),
       data = na.omit(movies2))
as_full.bas
```

According to this model, there is a 100% chance that ```imdb_rating``` will be included in the final model. Other noteworthy variables are ```runtime``` (~47%), ```critics_score``` (~89%). The variable with the nearest score to these is ```mpaa_rating_R:yes``` at ~20%.

```{r}
confint(coef(as_full.bas))
```


```{r}
summary(as_full.bas)
```

The best model chosen contains the variables runtime, imdb_rating and critics_score. Notice that this is the same model created by the backwards stepwise BIC method above.

Below, we can visualize the goodness of each of the models analyzed using the bas.lm function. The best model (rank 1) shows on the left, with the colored squares representing variables that would be selected for that particular model.

```{r, fig.width = 7, fig.height = 7}
image(as_full.bas, rotate = F)
```

```{r}
qqnorm(BIC.lm$residuals, col="red")
qqline(BIC.lm$residuals)
```

We observe a normal distribution here.

```{r}
plot(BIC.lm$residuals ~ BIC.lm$fitted, col="red")
abline(h=0, lty=2)
```

We notice some left-skewness here, but the data is generally scattered around 0.

```{r}
plot(abs(BIC.lm$residuals) ~ BIC.lm$fitted, col="red")
```

Here We do not meet the necessary condition.


### Part 5: Prediction

I am going to take the movie called "Doctor Strange" which was released in 2016 ( a movie that is not presented in the dataset). We extracted the information from the IMDB and Rotten tomatoes websites.

```{r}
doctor_strange <- data.frame(imdb_rating = 7.5, runtime = 97, critics_score = 40, mpaa_rating_R="no", thtr_rel_year=2016, best_pic_nom="no",best_actor_win="no", best_actress_win="no")
```

```{r}
predict(BIC.lm, newdata = doctor_strange, interval = "prediction", level = 0.95)
```

The BIC model predicts a score of 80.48538.

```{r}
predict(AIC.lm, newdata = doctor_strange, interval = "prediction", level = 0.95)
```

The AIC model predicts a score of 80.41053.

I have used both the BIC and AIC method. As the true score was 80, the BIC model was only marginally more accurate. 


### Part 6: Conclusion

The relationship between audience score and the variables in movies dataset was analyzed in project. We observed that there is a strong correlation between the audience_score with critics_score and imdb_rating. Using the AIC and BIC models we predicted the score for movie "Doctor Strange", in both we got the score almost similar. Even other factors may cause indirectly to the popularity of the movie.

